Hypothesis tests for multiple responses regression: effect of probiotics on addiction and binge eating disorder.

### 2023-10-19 5_artigos/the-international-journal-of-biostatistics/draft/ijb-article-lineu.tex
adiciona artigo ijb

% sage_latex_guidelines.tex V1.20, 14 January 2017

\documentclass[Review,sagev,times, doublespace]{sagej}

\usepackage{moreverb,url}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2016}

@@

\runninghead{Freitas et al}

\title{Hypothesis tests for multiple responses regression: effect of probiotics on addiction and binge eating disorder.}

\author{Lineu Alberto Cavazani de Freitas\affilnum{1}, Ligia de Oliveira Carlos\affilnum{2}, Antônio Carlos Ligocki Campos\affilnum{2} and Wagner Hugo Bonat\affilnum{3}}

\affiliation{
    \affilnum{1}Department of Informatics, Paraná Federal University, Paraná, Brazil\\
    \affilnum{2}Department of Surgical Clinic, Paraná Federal University, Paraná, Brazil\\
    \affilnum{3}Department of Statistics, Paraná Federal University, Paraná, Brazil\\    
}

\corrauth{Lineu Alberto Cavazani de Freitas, Department of Informatics, Paraná Federal University, R. Evaristo F. Ferreira da Costa, 383-391, Jardim das Américas, Curitiba, Paraná, 82590-300, Brazil.}

\email{lineuacf@gmail.com}

\begin{abstract}

Clinical trials are common in medical research where multiple non-Gaussian responses and time-dependent observations are frequent.
The analysis of data from these studies requires statistical modeling techniques that take these characteristics into account. 
We propose a general strategy based on the Wald statistics to perform hypothesis tests like ANOVAs, MANOVAs and multiple comparison tests on 
regression and dispersion parameters of multivariate covariance generalized linear models (McGLMs). 
McGLMs provide a general statistical modeling framework for normal and non-normal multivariate data analysis along with a wide range of correlation structures. 
We design different simulation scenarios to verify the properties of the proposed tests. 
The results are promising showing that the proposed tests present the levels of confidence close to the specified one for all simulation study scenarios.
Complementary to the proposal, we developed implementations in the R language to carry out the tests presented, the codes are available in the supplementary material.
The proposal is motivated by the analysis of a clinical trial that aims to evaluate the effect of the use of probiotics in the control of addiction and binge eating disorder in patients undergoing bariatric surgery. 
The subjects were separated into two groups (placebo and treatment) and evaluated at three different times. 
    The results indicate that addiction and binge eating disorder reduce over time, but there is no difference between groups at each time point.

\end{abstract}

\keywords{McGLM, Hypothesis tests, Wald test, ANOVA, MANOVA, Multiple comparisons}

@@

We present the dataset in \autoref{sec2} 2. In \autoref{sec3} 3 we present the review of the general structure and estimation of the parameters of a McGLM, based on the ideas of Bonat and Jørgensen \cite{Bonat16}. In \autoref{sec4} 4 our proposal is presented with the details of the Wald test to evaluate assumptions about parameters of a McGLM. In \autoref{sec5} 5 we present the results of the performance evaluation of the proposed test based on a simulation study. In the \autoref{sec6} 6 we apply the model and present the main results of the analysis of the data from the clinical study that aims to evaluate the use of probiotics in the control of addiction and binge eating disorder in patients undergoing bariatric surgery. Finally, the main results are discussed in \autoref{sec7} 7, including some directions for future investigations. The data sets and R codes are available online and can be assessed in \href{https://github.com/lineu96/paper-htmcglm}{https://github.com/lineu96/paper-htmcglm}.

@@

\begin{figure}[h]
\centering
\includegraphics[width=29.6pc,height=22.2pc]{FIGURE1.jpeg}
\caption{Exploratory graphic analysis: (a) YFAS histogram, (b) YFAS boxplots as a function of group, (c) YFAS boxplots as a function of moment, (d) BES histogram, (b) BES boxplots as a function of group, (c) BES boxplots as a function of moment. The asterisk in the boxplots indicates the mean.\label{fig1}}
\end{figure}

@@

\begin{table}[h]
\centering
\begin{tabular}{ccccc}
\hline
\multirow{2}{*}{Group} & \multirow{2}{*}{Moment}  & \multirow{2}{*}{n} & YFAS                  & BES                   \\ \cline{4-5} 
                       &                          &                    & Mean (standard deviation) & Mean (standard deviation) \\ \hline
Placebo                & T0                       & 33                 & 0.37 (0.26)           & 0.24 (0.20)           \\
Placebo                & T1                       & 32                 & 0.11 (0.15)           & 0.09 (0.10)           \\
Placebo                & T2                       & 22                 & 0.16 (0.15)           & 0.10 (0.12)           \\
Probiotic              & T0                       & 38                 & 0.49 (0.24)           & 0.32 (0.18)           \\
Probiotic              & T1                       & 37                 & 0.09 (0.12)           & 0.10 (0.08)           \\
Probiotic              & T2                       & 22                 & 0.10 (0.14)           & 0.07 (0.09)           \\ \hline
\end{tabular}
\caption{Number of individuals, mean and standard deviation for YFAS and BES for each combination of group and time point.}
\label{tab:tab1}
\end{table}

@@

\begin{table}[h]
\centering
\begin{tabular}{ccc}
\hline
Distribution & Link function     & Variance function \\ \hline
Normal       & Identity          & Constant           \\
Poisson      & Logarithmic       & Tweedie             \\
Bernoulli    & Logit             & Binomial            \\ \hline
\end{tabular}
\caption{Link and variance functions used in the models for each simulated distribution.}
\label{tab2}
\end{table}

@@

\begin{figure}[h]
\centering
\includegraphics[width=29.6pc,height=22.2pc]{FIGURE2.jpeg}
\caption{Simulation study results for the regression parameters.\label{fig2}}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=29.6pc,height=22.2pc]{FIGURE3.jpeg}
\caption{Simulation study results for the dispersion parameters.\label{fig3}}
\end{figure}

@@

In this application, for composing the linear predictor, we considered the fixed effects of moment (T0,T1,T2) and group (placebo, probiotc). Additionally, the effect of the interaction between these two explanatory variables was included in the model.

@@

\begin{figure}[h]
\centering
\includegraphics[width=29.4pc,height=12.6pc]{FIGURE4.jpeg}
\caption{Histogram of Pearson residuals by response.\label{fig4}}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=29.4pc,height=12.6pc]{FIGURE5.jpeg}
\caption{Pearson residuals versus predicted plot with smooth trend line for each response.\label{fig5}}
\end{figure}

@@

\begin{table}[h]
\centering
\begin{tabular}{c|cccccc}
\hline
\multirow{2}{*}{Parameter} & \multicolumn{3}{c}{YFAS}                                                                                             & \multicolumn{3}{c}{BES}                                                                         \\ \cline{2-7} 
                           & Estimate & \begin{tabular}[c]{@{}c@{}}Confidence \\ interval\end{tabular} & \multicolumn{1}{c|}{p-value}        & Estimate & \begin{tabular}[c]{@{}c@{}}Confidence \\ interval\end{tabular} & p-value        \\ \hline
$\beta_0$                  & -0.54      & (-0.87;-0.22)                                                     & \multicolumn{1}{c|}{\textless 0.01} & -1.13      & (-1.44;-0.83)                                                     & \textless 0.01 \\
$\beta_1$                  & -1.55      & (-2.17;-0.94)                                                     & \multicolumn{1}{c|}{\textless 0.01} & -1.16      & (-1.62;-0.69)                                                     & \textless 0.01 \\
$\beta_2$                  & -1.13      & (-1.75;-0.51)                                                     & \multicolumn{1}{c|}{\textless 0.01} & -1.05      & (-1.58;-0.52)                                                     & \textless 0.01 \\
$\beta_3$                  & 0.49       & (0.05;0.93)                                                       & \multicolumn{1}{c|}{0.0284} & 0.37       & (-0.03;0.77)                                                      & 0.0733           \\
$\beta_4$                  & -0.73      & (-1.60;0.14)                                                      & \multicolumn{1}{c|}{0.0}            & -0.33      & (-0.96;0.30)                                                      & 0.3081           \\
$\beta_5$                  & -0.98      & (-1.93;-0.03)                                                     & \multicolumn{1}{c|}{0.0429} & -0.80      & (-1.58;-0.02)                                                     & 0.0449 \\
$\tau_0$                   & 0.18       & (0.01;0.35)                                                       & \multicolumn{1}{c|}{0.0411} & 0.17       & (0.00;0.34)                                                       & 0.0458           \\
$\tau1$                    & 0.01       & (-0.02;0.04)                                                      & \multicolumn{1}{c|}{0.5718}           & 0.04       & (-0.01;0.10)                                                      & 0.1357           \\
$p$                        & 0.91       & (0.47;1.34)                                                       & \multicolumn{1}{c|}{\textless 0.05} & 1.23       & (0.77;1.68)                                                       & \textless 0.05 \\ \hline
\end{tabular}
\caption{Parameter estimates, 95\% confidence intervals and p-values of the model.}
\label{tab3}
\end{table}

\begin{figure}[h]
\centering
\includegraphics[width=31.5pc,height=18pc]{FIGURE6.jpeg}
\caption{Graph of model predicted for each time and group combination.\label{fig6}}
\end{figure}

@@

We can opt for a type II multivariate analysis of variance to assess the importance of the variables in the problem. This result, presented in the table \autoref{tab4}, points to the significant existence of the moment effect and the absence of the group effect, indicating that for both responses the metrics change over time but without change between groups.

\begin{table}[h]
\centering
\begin{tabular}{cccc}
\hline
Variable      & Degrees of freedom & Chi-square   & p-value        \\ \hline
Intercept     & 2                  & 53.1581      & \textless 0.01 \\
Moment        & 8                  & 139.0161     & \textless 0.01 \\
Group         & 6                  & 8.4928       & 0.2042         \\
Moment*Group  & 4                  & 6.9923       & 0.1363         \\ \hline
\end{tabular}
\caption{Type II multivariate analysis of variance.}
\label{tab4}
\end{table}

@@

\begin{table}[h]
\centering
\begin{tabular}{c|c|cc|cc}
\hline
              &                    & \multicolumn{2}{c|}{YFAS}     & \multicolumn{2}{c}{BES}       \\ \hline
Variable      & Degrees of freedom & Chi-square   & p-value        & Chi-square   & p-value        \\ \hline
Intercept     & 1                  & 10.6128      & \textless 0.01 & 53.1473      & \textless 0.01 \\
Moment        & 4                  & 102.9875     & \textless 0.01 & 99.5681      & \textless 0.01 \\
Group         & 3                  & 6.6837       & 0.0827         & 5.3083       & 0.1506         \\
Moment*Group  & 2                  & 5.5984       & 0.0609         & 4.2477       & 0.1196         \\ \hline
\end{tabular}
\caption{Type II univariate analysis of variance.}
\label{tab5}
\end{table}

As the analysis of variances pointed to a significant effect of categorical variables, we can explore which levels differ from each other. \autoref{tab6} shows two-by-two comparisons between moments. The results show that, for both responses, there are differences between the first versus second and first versus third moments, but the last two moments do not differ from each other. The table \autoref{tab7} presents the comparisons between groups for each time point for both responses. The results point to the absence of difference between groups at each moment.

\begin{table}[h]
\centering
\begin{tabular}{cccc}
\hline
Contrast  & Degrees of freedom & Chi-square   & p-value        \\ \hline
T0-T1     & 2                  & 97.9874      & \textless 0.01 \\
T0-T2     & 2                  & 67.2462      & \textless 0.01 \\
T1-T2     & 2                  & 2.4730       & 0.8712         \\ \hline
\end{tabular}
\caption{Two-by-two comparisons between moments for both responses.}
\label{tab6}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{cccc}
\hline
Contrast                 & Degrees of freedom & Chi-square   & p-value \\ \hline
T0:Placebo-T0:Probiotic  & 2                  & 5.5819       & 0.9204  \\
T1:Placebo-T1:Probiotic  & 2                  & 0.6096       & 1       \\
T2:Placebo-T2:Probiotic  & 2                  & 1.7645       & 1       \\ \hline
\end{tabular}
\caption{Two-by-two comparisons between groups for each time point for both responses.}
\label{tab7}
\end{table}

In the model, we include the information that there are measurements that were taken on the same individual. This information is declared in the matrix predictor that estimates a dispersion parameter associated with the matrix that indicates the relationship between individuals. A hypothesis of interest may be to assess whether there is evidence to believe that, in this problem, the measurements taken on the same individual are in fact correlated. For this, we can postulate hypotheses about the dispersion parameters. As with analyzes of variance, this can be done per response or for both responses simultaneously. The table \autoref{tab8} presents the results of the multivariate test, that is, it evaluates the hypothesis that in both responses the measures are correlated. The results indicate that there is no evidence to believe that the measures taken in the same individual are correlated.

\begin{table}[h]
\centering
\begin{tabular}{cccc}
\hline
Variable               & Degrees of freedom & Chi-square & p-value        \\ \hline
$\tau_0$ & 2                  & 7.1936       & 0.0274  \\
$\tau_1$ & 2                  & 2.3201       & 0.3135         \\ \hline
\end{tabular}
\caption{Type III multivariate analysis of variance for dispersion parameters.}
\label{tab8}
\end{table}

@@

\begin{acks}
The authors thank the reviewers for their constructive and helpful comments, which greatly improved the article.
\end{acks}

%\begin{biog}
%To typeset an
%  "Author biography" section.
%\end{biog}

%\begin{biogs}
%To typeset an
%  "Author Biographies" section.
%\end{biogs}

\begin{dci}
The Authors declares that there is no conflict of interest.
\end{dci}

\begin{funding}
This study was
financed in part by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior – Brasil (CAPES) – Finance Code 001.
\end{funding}

\begin{sm}
R code and data set are available at https://github.com/lineu96/paper-htmcglm
\end{sm}

\bibliographystyle{SageV}
\bibliography{draft-sm-lineu.bib}

@@

\begin{table}[h]
\centering
\begin{tabular}{c|cccc}
\hline
Null hypothesis & $\beta_0$ & $\beta_1$ & $\beta_2$ & $\beta_3$ \\ \hline
$H_{01}$      & 5         & 0         & 0         & 0         \\
$H_{02}$      & 4.85      & 0.05      & 0.05      & 0.05      \\
$H_{03}$      & 4.7       & 0.1       & 0.1       & 0.1       \\
$H_{04}$      & 4.55      & 0.15      & 0.15      & 0.15      \\
$H_{05}$      & 4.4       & 0.2       & 0.2       & 0.2       \\
$H_{06}$      & 4.25      & 0.25      & 0.25      & 0.25      \\
$H_{07}$      & 4.1       & 0.3       & 0.3       & 0.3       \\
$H_{08}$      & 3.95      & 0.35      & 0.35      & 0.35      \\
$H_{09}$      & 3.8       & 0.4       & 0.4       & 0.4       \\
$H_{10}$      & 3.65      & 0.45      & 0.45      & 0.45      \\
$H_{11}$      & 3.5       & 0.5       & 0.5       & 0.5       \\
$H_{12}$      & 3.35      & 0.55      & 0.55      & 0.55      \\
$H_{13}$      & 3.2       & 0.6       & 0.6       & 0.6       \\
$H_{14}$      & 3.05      & 0.65      & 0.65      & 0.65      \\
$H_{15}$      & 2.9       & 0.7       & 0.7       & 0.7       \\
$H_{16}$      & 2.75      & 0.75      & 0.75      & 0.75      \\
$H_{17}$      & 2.6       & 0.8       & 0.8       & 0.8       \\
$H_{18}$      & 2.45      & 0.85      & 0.85      & 0.85      \\
$H_{19}$      & 2.3       & 0.9       & 0.9       & 0.9       \\
$H_{20}$      & 2.15      & 0.95      & 0.95      & 0.95      \\ \hline
\end{tabular}
\caption{Tested hypotheses for regression parameters in models with response following Normal distribution.}
\label{tab:hipoteses_beta_normal}
\end{table}

>>>>

\documentclass[USenglish]{article}	
% for 2-column layout use \documentclass[USenglish,twocolumn]{article}

\usepackage[utf8]{inputenc}				%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}[2017/03/31]%(only for the luatex or the xetex engine)
\usepackage[big,online]{dgruyter}	%values: small,big | online,print,work
\usepackage{lmodern} 
\usepackage{microtype}
\usepackage[numbers,square,sort&compress]{natbib}

\usepackage{float}
\usepackage{anyfontsize}
\usepackage{multirow}

% New theorem-like environments will be introduced by using the commands \theoremstyle and \newtheorem.
% Please note that the environments proof and definition are already defined within dgryuter.sty.
\theoremstyle{dgthm}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{assertion}{Assertion}
\newtheorem{result}{Result}
\newtheorem{conclusion}{Conclusion}

\theoremstyle{dgdef}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}

@@

%%%--------------------------------------------%%%
	\articletype{Research Article}
	\received{Month	DD, YYYY}
	\revised{Month	DD, YYYY}
  \accepted{Month	DD, YYYY}
  \journalname{De~Gruyter~Journal}
  \journalyear{YYYY}
  \journalvolume{XX}
  \journalissue{X}
  \startpage{1}
  \aop
  \DOI{10.1515/sample-YYYY-XXXX}
%%%--------------------------------------------%%%

\title{Hypothesis tests for multiple responses regression: effect of probiotics on addiction and binge eating disorder.}
\runningtitle{Hypothesis tests for multiple responses regression.}
%\subtitle{Insert subtitle if needed}

\author*[1]{Lineu Alberto Cavazani de Freitas}
%\ use * to mark the author as the corresponding author
\author[2]{Ligia de Oliveira Carlos}
\author[2]{Antônio Carlos Ligocki Campos} 
\author[1]{Wagner Hugo Bonat}

\runningauthor{de Freitas et al.}

\affil[1]{\protect\raggedright 
Paraná Federal University, Department of Statistics, Curitiba, Brazil, e-mail: lineuacf@gmail.com, wbonat@ufpr.br}

\affil[2]{\protect\raggedright 
Paraná Federal University, Department of Surgical Clinic, Curitiba, Brazil, e-mail: ligia.oc@hotmail.com, aclcampos@hotmail.com}

%\communicated{...}
%\dedication{...}

\abstract{Clinical trials are common in medical research where multiple non-Gaussian responses and time-dependent observations are frequent.
The analysis of data from these studies requires statistical modeling techniques that take these characteristics into account. 
We propose a general strategy based on the Wald statistics to perform hypothesis tests like ANOVAs, MANOVAs and multiple comparison tests on 
regression and dispersion parameters of multivariate covariance generalized linear models (McGLMs). 
McGLMs provide a general statistical modeling framework for normal and non-normal multivariate data analysis along with a wide range of correlation structures. 
We design different simulation scenarios to verify the properties of the proposed tests. 
The results are promising showing that the proposed tests present the levels of confidence close to the specified one for all simulation study scenarios.
Complementary to the proposal, we developed implementations in the R language to carry out the tests presented, the codes are available in the supplementary material.
The proposal is motivated by the analysis of a clinical trial that aims to evaluate the effect of the use of probiotics in the control of addiction and binge eating disorder in patients undergoing bariatric surgery. 
The subjects were separated into two groups (placebo and treatment) and evaluated at three different times. 
    The results indicate that addiction and binge eating disorder reduce over time, but there is no difference between groups at each time point.}

\keywords{McGLM, Hypothesis tests, Wald test, ANOVA, MANOVA, Multiple comparisons.}

@@

We present the dataset in \autoref{sec2}. In \autoref{sec3} we present the review of the general structure and estimation of the parameters of a McGLM, based on the ideas of Bonat and Jørgensen \cite{Bonat16}. In \autoref{sec4} our proposal is presented with the details of the Wald test to evaluate assumptions about parameters of a McGLM. In \autoref{sec5} we present the results of the performance evaluation of the proposed test based on a simulation study. In the \autoref{sec6} we apply the model and present the main results of the analysis of the data from the clinical study that aims to evaluate the use of probiotics in the control of addiction and binge eating disorder in patients undergoing bariatric surgery. Finally, the main results are discussed in \autoref{sec7}, including some directions for future investigations. The data sets and R codes are available online and can be assessed in \href{https://github.com/lineu96/paper-htmcglm}{https://github.com/lineu96/paper-htmcglm}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=29.6pc,height=22.2pc]{FIGURES/FIGURE1.jpeg}
\caption{Exploratory graphic analysis: (a) YFAS histogram, (b) YFAS boxplots as a function of group, (c) YFAS boxplots as a function of moment, (d) BES histogram, (e) BES boxplots as a function of group, (f) BES boxplots as a function of moment. The asterisk in the boxplots indicates the mean.\label{fig1}}
\end{figure}

@@

\begin{table}[H]
\centering
\begin{tabular}{ccccc}
\hline
\multirow{2}{*}{Group} & \multirow{2}{*}{Moment}  & \multirow{2}{*}{n} & YFAS                  & BES                   \\ \cline{4-5} 
                       &                          &                    & Mean (standard deviation) & Mean (standard deviation) \\ \hline
Placebo                & T0                       & 33                 & 0.37 (0.26)           & 0.24 (0.20)           \\
Placebo                & T1                       & 32                 & 0.11 (0.15)           & 0.09 (0.10)           \\
Placebo                & T2                       & 22                 & 0.16 (0.15)           & 0.10 (0.12)           \\
Probiotic              & T0                       & 38                 & 0.49 (0.24)           & 0.32 (0.18)           \\
Probiotic              & T1                       & 37                 & 0.09 (0.12)           & 0.10 (0.08)           \\
Probiotic              & T2                       & 22                 & 0.10 (0.14)           & 0.07 (0.09)           \\ \hline
\end{tabular}
\caption{Number of individuals, mean and standard deviation for YFAS and BES for each combination of group and time point.}
\label{tab:tab1}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{ccc}
\hline
Distribution & Link function     & Variance function \\ \hline
Normal       & Identity          & Constant           \\
Poisson      & Logarithmic       & Tweedie             \\
Bernoulli    & Logit             & Binomial            \\ \hline
\end{tabular}
\caption{Link and variance functions used in the models for each simulated distribution.}
\label{tab2}
\end{table}

@@

\begin{figure}[H]
\centering
\includegraphics[width=29.6pc,height=22.2pc]{FIGURES/FIGURE2.jpeg}
\caption{Simulation study results for the regression parameters.\label{fig2}}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=29.6pc,height=22.2pc]{FIGURES/FIGURE3.jpeg}
\caption{Simulation study results for the dispersion parameters.\label{fig3}}
\end{figure}

@@

In this application, for composing the linear predictor, we considered the fixed effects of moment (T0,T1,T2) and group (placebo, probiotic). Additionally, the effect of the interaction between these two explanatory variables was included in the model.

@@

\begin{figure}[H]
\centering
\includegraphics[width=29.4pc,height=12.6pc]{FIGURES/FIGURE4.jpeg}
\caption{Histogram of Pearson residuals by response.\label{fig4}}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=29.4pc,height=12.6pc]{FIGURES/FIGURE5.jpeg}
\caption{Pearson residuals versus predicted plot with smooth trend line for each response.\label{fig5}}
\end{figure}

@@

\begin{table}[H]
\centering
\begin{tabular}{c|cccccc}
\hline
\multirow{2}{*}{Parameter} & \multicolumn{3}{c}{YFAS}                                                                                             & \multicolumn{3}{c}{BES}                                                                         \\ \cline{2-7}
                           & Estimate & \begin{tabular}[c]{@{}c@{}}Confidence \\ interval\end{tabular} & \multicolumn{1}{c|}{p-value}        & Estimate & \begin{tabular}[c]{@{}c@{}}Confidence \\ interval\end{tabular} & p-value        \\ \hline
$\beta_0$                  & -0.54      & (-0.87;-0.22)                                                     & \multicolumn{1}{c|}{\textless 0.01} & -1.13      & (-1.44;-0.83)                                                     & \textless 0.01 \\
$\beta_1$                  & -1.55      & (-2.17;-0.94)                                                     & \multicolumn{1}{c|}{\textless 0.01} & -1.16      & (-1.62;-0.69)                                                     & \textless 0.01 \\
$\beta_2$                  & -1.13      & (-1.75;-0.51)                                                     & \multicolumn{1}{c|}{\textless 0.01} & -1.05      & (-1.58;-0.52)                                                     & \textless 0.01 \\
$\beta_3$                  & 0.49       & (0.05;0.93)                                                       & \multicolumn{1}{c|}{0.0284} & 0.37       & (-0.03;0.77)                                                      & 0.0733           \\
$\beta_4$                  & -0.73      & (-1.60;0.14)                                                      & \multicolumn{1}{c|}{0.0}            & -0.33      & (-0.96;0.30)                                                      & 0.3081           \\
$\beta_5$                  & -0.98      & (-1.93;-0.03)                                                     & \multicolumn{1}{c|}{0.0429} & -0.80      & (-1.58;-0.02)                                                     & 0.0449 \\
$\tau_0$                   & 0.18       & (0.01;0.35)                                                       & \multicolumn{1}{c|}{0.0411} & 0.17       & (0.00;0.34)                                                       & 0.0458           \\
$\tau1$                    & 0.01       & (-0.02;0.04)                                                      & \multicolumn{1}{c|}{0.5718}           & 0.04       & (-0.01;0.10)                                                      & 0.1357           \\
$p$                        & 0.91       & (0.47;1.34)                                                       & \multicolumn{1}{c|}{\textless 0.05} & 1.23       & (0.77;1.68)                                                       & \textless 0.05 \\ \hline
\end{tabular}
\caption{Parameter estimates, 95\% confidence intervals and p-values of the model.}
\label{tab3}
\end{table}

\begin{figure}[H]
\centering
\includegraphics[width=31.5pc,height=18pc]{FIGURES/FIGURE6.jpeg}
\caption{Graph of model predicted for each time and group combination.\label{fig6}}
\end{figure}

@@

We can opt for a type II multivariate analysis of variance to assess the importance of the variables in the problem. This result, presented in the \autoref{tab4}, points to the significant existence of the moment effect and the absence of the group effect, indicating that for both responses the metrics change over time but without change between groups.

\begin{table}[H]
\centering
\begin{tabular}{cccc}
\hline
Variable      & Degrees of freedom & Chi-square   & p-value        \\ \hline
Intercept     & 2                  & 53.1581      & \textless 0.01 \\
Moment        & 8                  & 139.0161     & \textless 0.01 \\
Group         & 6                  & 8.4928       & 0.2042         \\
Moment*Group  & 4                  & 6.9923       & 0.1363         \\ \hline
\end{tabular}
\caption{Type II multivariate analysis of variance.}
\label{tab4}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{c|c|cc|cc}
\hline
              &                    & \multicolumn{2}{c|}{YFAS}     & \multicolumn{2}{c}{BES}       \\ \hline
Variable      & Degrees of freedom & Chi-square   & p-value        & Chi-square   & p-value        \\ \hline
Intercept     & 1                  & 10.6128      & \textless 0.01 & 53.1473      & \textless 0.01 \\
Moment        & 4                  & 102.9875     & \textless 0.01 & 99.5681      & \textless 0.01 \\
Group         & 3                  & 6.6837       & 0.0827         & 5.3083       & 0.1506         \\
Moment*Group  & 2                  & 5.5984       & 0.0609         & 4.2477       & 0.1196         \\ \hline
\end{tabular}
\caption{Type II univariate analysis of variance.}
\label{tab5}
\end{table}

As the analysis of variances pointed to a significant effect of categorical variables, we can explore which levels differ from each other. \autoref{tab6} shows two-by-two comparisons between moments. The results show that, for both responses, there are differences between the first versus second and first versus third moments, but the last two moments do not differ from each other. The \autoref{tab7} presents the comparisons between groups for each time point for both responses. The results point to the absence of difference between groups at each moment.

\begin{table}[H]
\centering
\begin{tabular}{cccc}
\hline
Contrast  & Degrees of freedom & Chi-square   & p-value        \\ \hline
T0-T1     & 2                  & 97.9874      & \textless 0.01 \\
T0-T2     & 2                  & 67.2462      & \textless 0.01 \\
T1-T2     & 2                  & 2.4730       & 0.8712         \\ \hline
\end{tabular}
\caption{Two-by-two comparisons between moments for both responses.}
\label{tab6}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{cccc}
\hline
Contrast                 & Degrees of freedom & Chi-square   & p-value \\ \hline
T0:Placebo-T0:Probiotic  & 2                  & 5.5819       & 0.9204  \\
T1:Placebo-T1:Probiotic  & 2                  & 0.6096       & 1       \\
T2:Placebo-T2:Probiotic  & 2                  & 1.7645       & 1       \\ \hline
\end{tabular}
\caption{Two-by-two comparisons between groups for each time point for both responses.}
\label{tab7}
\end{table}

In the model, we include the information that there are measurements that were taken on the same individual. This information is declared in the matrix predictor that estimates a dispersion parameter associated with the matrix that indicates the relationship between individuals. A hypothesis of interest may be to assess whether there is evidence to believe that, in this problem, the measurements taken on the same individual are in fact correlated. For this, we can postulate hypotheses about the dispersion parameters. As with analyzes of variance, this can be done per response or for both responses simultaneously. The \autoref{tab8} presents the results of the multivariate test, that is, it evaluates the hypothesis that in both responses the measures are correlated. The results indicate that there is no evidence to believe that the measures taken in the same individual are correlated.

\begin{table}[H]
\centering
\begin{tabular}{cccc}
\hline
Variable               & Degrees of freedom & Chi-square & p-value        \\ \hline
$\tau_0$ & 2                  & 7.1936       & 0.0274  \\
$\tau_1$ & 2                  & 2.3201       & 0.3135         \\ \hline
\end{tabular}
\caption{Type III multivariate analysis of variance for dispersion parameters.}
\label{tab8}
\end{table}

@@

\begin{acknowledgement}
The authors thank the reviewers for their constructive and helpful comments, which greatly improved the article.
\end{acknowledgement}

\begin{funding}
This study was financed in part by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior - Brasil (CAPES) -Finance Code 001.
\end{funding}

\bibliographystyle{unsrt}
\bibliography{bib.bib}

@@

\begin{table}[H]
\centering
\begin{tabular}{c|cccc}
\hline
Null hypothesis & $\beta_0$ & $\beta_1$ & $\beta_2$ & $\beta_3$ \\ \hline
$H_{01}$      & 5         & 0         & 0         & 0         \\
$H_{02}$      & 4.85      & 0.05      & 0.05      & 0.05      \\
$H_{03}$      & 4.7       & 0.1       & 0.1       & 0.1       \\
$H_{04}$      & 4.55      & 0.15      & 0.15      & 0.15      \\
$H_{05}$      & 4.4       & 0.2       & 0.2       & 0.2       \\
$H_{06}$      & 4.25      & 0.25      & 0.25      & 0.25      \\
$H_{07}$      & 4.1       & 0.3       & 0.3       & 0.3       \\
$H_{08}$      & 3.95      & 0.35      & 0.35      & 0.35      \\
$H_{09}$      & 3.8       & 0.4       & 0.4       & 0.4       \\
$H_{10}$      & 3.65      & 0.45      & 0.45      & 0.45      \\
$H_{11}$      & 3.5       & 0.5       & 0.5       & 0.5       \\
$H_{12}$      & 3.35      & 0.55      & 0.55      & 0.55      \\
$H_{13}$      & 3.2       & 0.6       & 0.6       & 0.6       \\
$H_{14}$      & 3.05      & 0.65      & 0.65      & 0.65      \\
$H_{15}$      & 2.9       & 0.7       & 0.7       & 0.7       \\
$H_{16}$      & 2.75      & 0.75      & 0.75      & 0.75      \\
$H_{17}$      & 2.6       & 0.8       & 0.8       & 0.8       \\
$H_{18}$      & 2.45      & 0.85      & 0.85      & 0.85      \\
$H_{19}$      & 2.3       & 0.9       & 0.9       & 0.9       \\
$H_{20}$      & 2.15      & 0.95      & 0.95      & 0.95      \\ \hline
\end{tabular}
\caption{Tested hypotheses for regression parameters in models with response following Normal distribution.}
\label{tab:hipoteses_beta_normal}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{c|cccc}
\hline
Null hypothesis & $\beta_0$ & $\beta_1$ & $\beta_2$ & $\beta_3$ \\ \hline
$H_{01}$      & 2.3       & 0         & 0         & 0         \\
$H_{02}$      & 2.25      & 0.017     & 0.017     & 0.017     \\
$H_{03}$      & 2.2       & 0.033     & 0.033     & 0.033     \\
$H_{04}$      & 2.15      & 0.05      & 0.05      & 0.05      \\
$H_{05}$      & 2.10      & 0.067     & 0.067     & 0.067     \\
$H_{06}$      & 2.05      & 0.083     & 0.083     & 0.083     \\
$H_{07}$      & 2         & 0.1       & 0.1       & 0.1       \\
$H_{08}$      & 1.95      & 0.117     & 0.117     & 0.117     \\
$H_{09}$      & 1.9       & 0.133     & 0.133     & 0.133     \\
$H_{10}$      & 1.85      & 0.15      & 0.15      & 0.15      \\
$H_{11}$      & 1.8       & 0.167     & 0.167     & 0.167     \\
$H_{12}$      & 1.75      & 0.167     & 0.167     & 0.167     \\
$H_{13}$      & 1.7       & 0.2       & 0.2       & 0.2       \\
$H_{14}$      & 1.65      & 0.217     & 0.217     & 0.217     \\
$H_{15}$      & 1.6       & 0.233     & 0.233     & 0.233     \\
$H_{16}$      & 1.55      & 0.25      & 0.25      & 0.25      \\
$H_{17}$      & 1.5       & 0.267     & 0.267     & 0.267     \\
$H_{18}$      & 1.45      & 0.283     & 0.283     & 0.283     \\
$H_{19}$      & 1.4       & 0.3       & 0.3       & 0.3       \\
$H_{20}$      & 1.35      & 0.317     & 0.317     & 0.317      \\ \hline
\end{tabular}
\caption{Tested hypotheses for regression parameters in models with response following Poisson distribution.}
\label{tab:hipoteses_beta_poisson}
\end{table}
